# Interactive component

##Interactive
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(plotly)
library(rCharts)
```

```{r}
data<-read.csv(file = 'ca-educational-attainment-personal-income-2008-2014.csv', header = TRUE)
data_p<-data %>% filter(Age == '18 to 64')
data_p$Educational.Attainment<-factor(data_p$Educational.Attainment,
                                      levels = c("Bachelor's degree or higher","Some college, less than 4-yr degree",
                                                 "High school or equivalent","No high school diploma"))
data_p$Personal.Income<-factor(data_p$Personal.Income, 
                               levels = c('$75,000 and over','$50,000 to $74,999','$35,000 to $49,999','$25,000 to $34,999',
                                          '$15,000 to $24,999','$10,000 to $14,999','$5,000 to $9,999','No Income'))
data_p$Year<-fct_recode(data_p$Year,
                        "2008"='01/01/2008 12:00:00 AM',"2009"='01/01/2009 12:00:00 AM',
                        "2010"='01/01/2010 12:00:00 AM',"2011"='01/01/2011 12:00:00 AM',
                        "2012"='01/01/2012 12:00:00 AM',"2013"='01/01/2013 12:00:00 AM',
                        "2014"='01/01/2014 12:00:00 AM')

data_p$Educational.Attainment<- fct_recode(data_p$Educational.Attainment,
                                               ">=Bachelor" = "Bachelor's degree or higher","College (<4year)" = "Some college, less than 4-yr degree",
                                               "High school or equivalent" = "High school or equivalent","< high school" = "No high school diploma")
data_p$Personal.Income<- fct_recode(data_p$Personal.Income,
                                        "[75k,inf)"='$75,000 and over',"[50k,75k)"='$50,000 to $74,999',"[35k,50k)"='$35,000 to $49,999',
                                        "[25k,35k)"='$25,000 to $34,999',"[15k,25k)"='$15,000 to $24,999',"[10k,15k)"='$10,000 to $14,999',
                                        "[5k,10k)"='$5,000 to $9,999','No Income'='No Income')
```


### line plot

population proportion for salary level for different year. 
interactive in showing the exact data when mouse is on points.
```{r}
data_inter<-data_p %>% group_by(Year, Personal.Income) %>%  #if we study gender, add Gender
  mutate(total = sum(Population.Count)) %>% group_by(Year) %>% 
  mutate(year_total = sum(Population.Count)) %>% ungroup() %>%
  summarise(Year = Year, Personal.Income = Personal.Income, ratio = total/year_total) %>% 
  unique()

fig1 <- ggplot(data_inter) + geom_point(aes(x = Year, y = ratio, color = Personal.Income))+
  geom_line(aes(x = Year, y = ratio, group = Personal.Income, color = Personal.Income))

ggplotly(fig1)

data_inter_gender<-data_p %>% group_by(Gender,Year, Personal.Income) %>%  #if we study gender, add Gender
  mutate(total = sum(Population.Count)) %>% group_by(Gender, Year) %>% 
  mutate(year_total_G = sum(Population.Count)) %>% ungroup() %>%
  summarise(Year = Year, Gender = Gender, Personal.Income = Personal.Income, ratio = total/year_total_G) %>% 
  unique()

fig2 <- ggplot(data_inter_gender) + geom_point(aes(x = Year, y = ratio, color = Personal.Income))+
  geom_line(aes(x = Year, y = ratio, group = Personal.Income, color = Personal.Income))+facet_grid(~Gender)

ggplotly(fig2)
```

### interactive choices for users, bar plots
This part is not compatible with the book??(perhaps we can find a way out) 
You can check this part in the latter section of interactive.R by running on your local Rstudio.
Perhaps We can realize the interactive plots by inserting a link?






